# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
setwd("/Users/rodolfofernandez/Desktop/RepData_PeerAssessment1")
act.data <- read.csv("activity.csv")
names(act.data)
head(act.data)
summary(act.data)
#process dates
library(lubridate)   
act.data$date <- ymd(act.data$date)
```
## What is mean total number of steps taken per day?
```{r}
hist(tapply(act.data$steps , act.data$date, sum, na.rm = T), col = "purple", xlab="Total steps per day", main = "Hist of total steps per day")
mean(tapply(act.data$steps , act.data$date, sum, na.rm = T), na.rm=T)
median(tapply(act.data$steps , act.data$date, sum, na.rm = T), na.rm=T)
```
## What is the average daily activity pattern?
```{r}
avg.ints <- tapply(act.data$steps , act.data$interval, mean, na.rm = T)
plot(as.numeric(rownames(avg.ints)), avg.ints, type = "l", lwd = 2, col = "blue", main = "Average interval steps across all days", xlab = "Interval")
```
## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(avg.ints[which.max(avg.ints)])
```
## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
length(which(is.na(act.data$steps == T) | is.na(act.data$date == T) | is.na(act.data$interval == T) ))
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
    The substitution that will be used to fill up all the NA's will be the average of the interval to which the NA belongs to (not rounded). It will be done along with the next point:
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
new.frame <- act.data
for(i in 1:length(act.data$steps)){
        if(is.na(new.frame[i, 1]) == T ){
                new.frame[i, 1] <- avg.ints[as.character(new.frame[i,3])]  
        }  
}
head(new.frame)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
hist(tapply(new.frame$steps , new.frame$date, sum, na.rm = T), col = "light blue", xlab="Total steps per day", main = "Hist of total steps per day")
mean(tapply(new.frame$steps , new.frame$date, sum, na.rm = T), na.rm=T)
median(tapply(new.frame$steps , new.frame$date, sum, na.rm = T), na.rm=T)
```

The mean increased, the median also, but not as much as the mean, also, the histogram seems a little bit more symmetric, and may have a normality appearance.

## Are there differences in activity patterns between weekdays and weekends?
